---
title: "Conductance"
author: "Kauanoe Greene"
date: "2024-11-14"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      fig.path = "../Output/")
```


# Libraries  

```{r}

# visuals
library(tidyverse)
library(dplyr)
library(tidytext)
library(here)

# models (stats)
library(lmerTest)
library (RLRsim)
library(Matrix)
# library(lme4)
library(lattice)
library(car)
library(nlme)

```

# Data  

```{r}
# Data uploading
functional.phase.data <- read_csv(here("Project", "Data", "functional.phase.data.csv"))
glimpse(functional.phase.data)
```

# Statistic Modeling: Chlorophyll Content Data  

- PT.I: Primary interaction treatment  
- PT.II: Reduced models  
- PT.III: Comparison models  

## PT.I: Full Model 
- main interaction  
- includes all main factors and interactions  
- F-statistic for a fixed effects mixed model  
- results table: drought column  

```{r}

# drought column result: (F-stat) 20.508, (p-val) 0.001138

cond.full.A = lmer(Conductance ~ Treatment + 
                    (1|ID) + 
                    (1|Population) + 
                    (1|Week) + 
                    (1|Population:Treatment) +  
                    (1|Treatment:Week), 
                  data = functional.phase.data, 
                  na.action = na.omit)

Anova(cond.full.A, test.statistic="F", type=3)
```

```{r}
# Normality check: untransformed data  
# Checking normality of residuals in untransformed data
# W = 0.99658, p-value = 0.01645

shapiro.test(resid(cond.full.A)) 
```

```{r}
# untransformed data: QQPlot  
qqnorm(resid(cond.full.A))
```

```{r}

# log-transformed data
# this will be used in the next steps when we look at it as log-transformed data.
log.cond <- log(functional.phase.data$Conductance)
glimpse(log.cond)
```

```{r}

# drought column result: (F-stat) 19.7, (p-val) 0.0009928

cond.full.B = lmer(log.cond ~ Treatment + 
                    (1|ID) + 
                    (1|Population) + 
                    (1|Week) + 
                    (1|Population:Treatment) +  
                    (1|Treatment:Week), 
                  data = functional.phase.data, 
                  na.action = na.omit)

Anova(cond.full.B, test.statistic="F", type=3)
```

```{r}

# transformed data: QQPlot 
qqnorm(resid(cond.full.B))
```

## Reduced Models  

- work bottom to top approach  
- drop the interaction  
- keep main factors  
- always keep “Treatment” and “(1|ID)” in all models  
- our goal is to find out if any of these factors improve model fit →   
- does it account for more of the variance in our response trait?  
- if it doesn’t, the chi square value will NOT be significant.  
- the factor we drop from the reduced model is the factor of focus. We want to know if adding this factor accounts for MORE variance than the treatment alone. A.K.A. we are investigating the effect of the factor we drop. 


```{r}

# 1: (Population:Treatment): chisq 124.56

cond.red.1 = lmer(Conductance ~ Treatment + 
                   (1|ID) + 
                   (1|Population) + 
                   (1|Week) + 
                   (1|Population:Treatment), 
                 data = functional.phase.data, 
                 na.action = na.omit)
Anova(cond.red.1)
```

```{r}

# 2: (Treatment:Time): chisq 23.594

cond.red.2 = lmer(Conductance ~ Treatment + 
                   (1|ID) + 
                   (1|Population) + 
                   (1|Week) + 
                   (1|Treatment:Week), 
                 data = functional.phase.data,  
                 na.action = na.omit)
Anova(cond.red.2)
```

```{r}

# 3: no interactions: chisq 360.78

cond.red.3 = lmer(Conductance ~ Treatment + 
                     (1|ID) + 
                     (1|Population) + 
                     (1|Week), 
                   data = functional.phase.data, 
                   na.action = na.omit)
Anova(cond.red.3)
```

```{r}

# 4: effect of week: chisq 360.39

cond.red.4 = lmer(Conductance ~ Treatment + 
                     (1|ID) + 
                     (1|Population), 
                   data = functional.phase.data, 
                   na.action = na.omit)
Anova(cond.red.4)
```

```{r}

# 5: effect of population: chisq 280.21

cond.red.5 = lmer(Conductance ~ Treatment + 
                         (1|ID) + 
                         (1|Week), 
                       data = functional.phase.data, 
                       na.action = na.omit)
Anova(cond.red.5)
```

## Comparison Models  
- the question: does adding the interaction/factors improve the model fit?
- comp.model.1 = anova(full.model, red.model.1) # test treatment x time interaction  
- comp.model.2 = anova(full.model, red.model.2) # test treatment x population interaction  
- comp.model.3 = anova(red.model.3, red.model.4) # test effect of week  
- comp.model.4= anova(red.model.3, red.model.5) # test effect of population  

```{r}

# 1: Test (treatment:time) interaction: chisq 74.006
anova(cond.full.A, cond.red.1)
```


```{r}

# 2A: Test (treatment:population) interaction: chisq 5.8752
anova(cond.full.A, cond.red.2)
```


```{r}

# 2B: Test (treatment:population) interaction: chisq 68.131
anova(cond.red.1, cond.red.2)
```

```{r}

# 3: Test the effect of week: chisq 43.709
anova(cond.red.3, cond.red.4)
```


```{r}

# 4: Test the effect of population: chisq 26.174
anova(cond.red.3, cond.red.5)
```

# Plot  

```{r}

# Data uploading
functional.phase.data <- read_csv(here("Project", "Data", "functional.phase.data.csv"))
glimpse(functional.phase.data)
view(functional.phase.data)
```


```{r}

# group control and treatment
# group by phase
# summarise mean

glimpse(functional.phase.data)

mean_data <- functional.phase.data %>% 
  select(Week, Population, Treatment, Chlorophyll, Conductance, ETRmax, Phase) %>% 
  group_by(Week, Population, Treatment, Phase) %>% 
  summarize(mean_chl = mean(Chlorophyll, na.rm = TRUE), 
            mean_cond = mean(Conductance, na.rm = TRUE), 
            mean_etrm = mean(ETRmax, na.rm = TRUE))

view(mean_data)
```


```{r}

# Chlorophyll content over time

condplot <- mean_data %>% # datasheet
  ggplot(aes(x = Week, # x-axis
             y = mean_cond, # y-axis
             color = Treatment)) + # colors
  geom_point() +  # data points
  geom_line() +  # plot
  labs(subtitle = "Detecting intraspecific variation in plasticity of rates of stomatal conductance in response to drought stress", # plot subtitle
       caption = "Data sourced from: Greene 2023", # plot caption
       x = "Time (Weeks)", # x-axis label
       y = "Mean Stomtal Conductance") + # y-axis label
  ggtitle("Effects of Drought on Stomatal Conductance in 'A'ali'i") + # plot title
  facet_wrap(~Population) + # create panels for each population!
  scale_color_manual(breaks= c("C", "PD"), labels = c("Control Group", "Pulse Drought Group"), values = c("darkgreen", "brown")) + # rename legend variables
  theme(plot.title = element_text(face = "bold", color = "black", hjust = 0.5), # bold title
        axis.text.x = element_text(size = 8), axis.title.x = element_text(size = 10), # adjust x-axis labels
        axis.text.y = element_text(size = 8), axis.title.y = element_text(size = 10),   # adjust y-axis labels
        legend.position = "top", 
        plot.subtitle = element_text(size = 10, hjust = 0.5), 
        legend.title = element_blank())

# save plot to my output folder
ggsave(here("Project", "Output", "condplot.png")) 

# view plot
condplot
```


```{r}

# group by phases

phase_data <- functional.phase.data %>% 
  select(Week, Population, Treatment, Chlorophyll, Conductance, ETRmax, Phase) %>% 
  group_by(Population, Treatment, Phase) %>% 
  summarize(mean_chl = mean(Chlorophyll, na.rm = TRUE), 
            mean_cond = mean(Conductance, na.rm = TRUE), 
            mean_etrm = mean(ETRmax, na.rm = TRUE))

view(phase_data)
```


```{r}

# plot recovery phases

condphase_plot <- phase_data %>% # datasheet
  ggplot(aes(x = Phase, # x-axis
             y = mean_cond, # y-axis
             color = Treatment)) + # colors
  geom_point() +  # data points # plot
  geom_line(aes(group = Treatment)) + 
  labs(subtitle = "Investigating the priming effects of drought treatment on stomatal conductance throughout recovery phases", # plot subtitle
       caption = "Data sourced from: Greene 2023", # plot caption
       x = "Recovery Phase", # x-axis label
       y = "Mean Stomatal Conductance") + # y-axis label
  ggtitle("Drought Priming Effect on Stomatal Conductance in 'A'ali'i") + # plot title
  facet_wrap(~Population) + # create panels for each population!
  scale_color_manual(breaks= c("C", "PD"), labels = c("Control Group", "Pulse Drought Group"), values = c("darkgreen", "brown")) + # rename legend variables
  scale_x_discrete(labels = c('Treatment', 'Pre', 'Mid', 'Post')) +
theme(plot.title = element_text(face = "bold", color = "black", hjust = 0.5), # bold title
        axis.text.x = element_text(size = 8), axis.title.x = element_text(size = 10), # adjust x-axis labels
        axis.text.y = element_text(size = 8), axis.title.y = element_text(size = 10),   # adjust y-axis labels
        legend.position = "top", 
        plot.subtitle = element_text(size = 10, hjust = 0.5), 
        legend.title = element_blank())

# save plot to my output folder
ggsave(here("Project", "Output", "condphase_plot.png")) 

# view plot
condphase_plot
```

