---
title: "Functional Trait Data!"
author: "Kauanoe Greene"
date: "2024-11-12"
output:
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      fig.path = "../Output/")
```


```{r Libraries}

# LIBRARIES

# visuals
library(tidyverse)
library(tidytext)
library(here)

# models (stats)
library(lmerTest)
library (RLRsim)
library(Matrix)
# library(lme4)
library(lattice)
library(car)
library(nlme)

```

```{r Data Upload}

functional.phase.data <- read_csv(here("Project", "Data", "functional.phase.data.csv"))
glimpse(functional.phase.data)

```
# Data  

```{r Data Analyses}

# DATA ANALYSES: Means

means <- functional.phase.data %>% 
  group_by(Population, Week, Treatment) %>% 
  summarise(mean.cond = mean(Conductance, na.rm = TRUE), 
            mean.chl = mean(Chlorophyll, na.rm = TRUE), 
            mean.alpha = mean(Alpha, na.rm = TRUE), 
            mean.etrm= mean(ETRmax, na.rm = TRUE), 
            mean.ek = mean(Ek, na.rm = TRUE), 
            mean.npq = mean(NPQmax, na.rm = TRUE), 
            mean.fvfm = mean(FvFm, na.rm = TRUE))

```

# Data Visualization  


```{r conductance plot}

# CONDUCTANCE PLOT
# Let us create a plot that will communicate our key takeaway
# Main question: how does drought affect the performance of plants across populations, measured via stomatal conductance?
# Key finding: stomatal conductance varies amongst control and drought treatment groups across populations over time.
# Stomatal conductance data was collected on a weekly basis

conductanceplot <- means %>% # datasheet
  ggplot(aes(x = Week, # x-axis
             y = mean.cond, # y-axis
             color = Treatment)) + # colors
  geom_point() +  # data points
  geom_line() +  # plot
  labs(subtitle = "Detecting intraspecific variation in plasticity of rates of stomatal conductance in response to drought stress", # plot subtitle
       caption = "Data sourced from: Greene 2023", # plot caption
       x = "Time (Weeks)", # x-axis label
       y = "Mean Stomatal Conductance (mmol m^(-2) s^(-1)") + # y-axis label
  ggtitle("Effects of Drought on Leaf Stomatal Conductance in 'A'ali'i") + # plot title
  facet_wrap(~Population) + # create panels for each population!
  scale_color_manual(breaks= c("C", "PD"), labels = c("Control Group", "Pulse Drought Group"), values = c("darkblue", "maroon")) + # rename legend variables
  theme(plot.title = element_text(face = "bold", color = "black", hjust = 0.5), # bold title
        axis.text.x = element_text(size = 8), axis.title.x = element_text(size = 10), # adjust x-axis labels
        axis.text.y = element_text(size = 8), axis.title.y = element_text(size = 10),   # adjust y-axis labels
        legend.position = "top", 
        plot.subtitle = element_text(size = 10, hjust = 0.5), 
        legend.title = element_blank())

# save plot to my output folder
ggsave(here("Project", "Output", "conductanceplot.png")) 

# view plot
conductanceplot
```

```{r chlorophyll plot}

# CHLOROPHYLL CONTENT

chlplot <- means %>% # datasheet
  ggplot(aes(x = Week, # x-axis
             y = mean.chl, # y-axis
             color = Treatment)) + # colors
  geom_point() +  # data points
  geom_line() +  # plot
  labs(subtitle = "Detecting intraspecific variation in plasticity of rates of chlorophyll content in response to drought stress", # plot subtitle
       caption = "Data sourced from: Greene 2023", # plot caption
       x = "Time (Weeks)", # x-axis label
       y = "Mean Chlorophyll Content") + # y-axis label
  ggtitle("Effects of Drought on Leaf Chlorophyll Content in 'A'ali'i") + # plot title
  facet_wrap(~Population) + # create panels for each population!
  scale_color_manual(breaks= c("C", "PD"), labels = c("Control Group", "Pulse Drought Group"), values = c("darkgreen", "brown")) + # rename legend variables
  theme(plot.title = element_text(face = "bold", color = "black", hjust = 0.5), # bold title
        axis.text.x = element_text(size = 8), axis.title.x = element_text(size = 10), # adjust x-axis labels
        axis.text.y = element_text(size = 8), axis.title.y = element_text(size = 10),   # adjust y-axis labels
        legend.position = "top", 
        plot.subtitle = element_text(size = 10, hjust = 0.5), 
        legend.title = element_blank())

# save plot to my output folder
ggsave(here("Project", "Output", "chlplot.png")) 

# view plot
chlplot
```
```{r fvfm plot}

# FV/FM Plot

fvfmplot <- means %>% # datasheet
  ggplot(aes(x = Week, # x-axis
             y = mean.fvfm, # y-axis
             color = Treatment)) + # colors
  geom_point() +  # data points
  geom_line() +  # plot
  labs(subtitle = "Detecting intraspecific variation in plasticity of photsynthetic yield in response to drought stress", # plot subtitle
       caption = "Data sourced from: Greene 2023", # plot caption
       x = "Time (Weeks)", # x-axis label
       y = "Mean Fv/Fm") + # y-axis label
  ggtitle("Effects of Drought on Leaf Stomatal Conductance in 'A'ali'i") + # plot title
  facet_wrap(~Population) + # create panels for each population!
  scale_color_manual(breaks= c("C", "PD"), labels = c("Control Group", "Pulse Drought Group"), values = c("orange", "maroon")) + # rename legend variables
  theme(plot.title = element_text(face = "bold", color = "black", hjust = 0.5), # bold title
        axis.text.x = element_text(size = 8), axis.title.x = element_text(size = 10), # adjust x-axis labels
        axis.text.y = element_text(size = 8), axis.title.y = element_text(size = 10),   # adjust y-axis labels
        legend.position = "top", 
        plot.subtitle = element_text(size = 10, hjust = 0.5), 
        legend.title = element_blank())

# save plot to my output folder
ggsave(here("Project", "Output", "fvfmplot.png")) 

# view plot
fvfmplot
```

```{r npq plot}

# NQPmax Plot

npqplot <- means %>% # datasheet
  ggplot(aes(x = Week, # x-axis
             y = mean.npq, # y-axis
             color = Treatment)) + # colors
  geom_point() +  # data points
  geom_line() +  # plot
  labs(subtitle = "Detecting intraspecific variation in plasticity of rates of maximum non-photochemical quenching values in response to drought stress", # plot subtitle
       caption = "Data sourced from: Greene 2023", # plot caption
       x = "Time (Weeks)", # x-axis label
       y = "Mean NPQ Max") + # y-axis label
  ggtitle("Effects of Drought on Leaf Stomatal Conductance in 'A'ali'i") + # plot title
  facet_wrap(~Population) + # create panels for each population!
  scale_color_manual(breaks= c("C", "PD"), labels = c("Control Group", "Pulse Drought Group"), values = c("violet", "maroon")) + # rename legend variables
  theme(plot.title = element_text(face = "bold", color = "black", hjust = 0.5), # bold title
        axis.text.x = element_text(size = 8), axis.title.x = element_text(size = 10), # adjust x-axis labels
        axis.text.y = element_text(size = 8), axis.title.y = element_text(size = 10),   # adjust y-axis labels
        legend.position = "top", 
        plot.subtitle = element_text(size = 10, hjust = 0.5), 
        legend.title = element_blank())

# save plot to my output folder
ggsave(here("Project", "Output", "npqplot.png")) 

# view plot
npqplot
```

# Statistic Modeling: Chlorophyll Content Data  

- PT.I: Primary interaction treatment  
- PT.II: Reduced models  
- PT.III: Comparison models


```{r}

# Data set: functional.phase.data
glimpse(functional.phase.data)

```


## PT.I: Full Model 
- main interaction  
- includes all main factors and interactions  
- F-statistic for a fixed effects mixed model  
- results table: drought column  


```{r Statistical Models: Chlorophyll Content}

# Drought column result: (F-stat) 3.9797, (p-val) 0.07637

chl.full.A = lmer(Chlorophyll ~ Treatment + 
                    (1|ID) + 
                    (1|Population) + 
                    (1|Week) + 
                    (1|Population:Treatment) +  
                    (1|Treatment:Week), 
                  data = functional.phase.data, 
                  na.action = na.omit)

Anova(chl.full.A, test.statistic="F", type=3)
```
```{r chl shapiro A}

# NORMALITY CHECK: UNTRANSFORMED  
# Checking normality of residuals in untransformed data
# W = 0.98162, p-value = 1.319e-10

shapiro.test(resid(chl.full.A)) 
```
```{r chl qqnorm A}

# Untransformed data: QQPlot  
qqnorm(resid(chl.full.A))

```
```{r chl log A}

# Log-transformed data
# This will be used in the next steps when we look at it as log-transformed data.

log.chl <- log(functional.phase.data$Chlorophyll)
glimpse(log.chl)

```

```{r chl primary B}

# Drought column result: (F-stat) 0.5701, (p-val) 0.4732

chl.full.B = lmer(log.chl ~ Treatment + 
                    (1|ID) + 
                    (1|Population) + 
                    (1|Week) + 
                    (1|Population:Treatment) +  
                    (1|Treatment:Week), 
                  data = functional.phase.data, 
                  na.action = na.omit)

Anova(chl.full.B, test.statistic="F", type=3)
```

```{r chl shapiro B}

# NORMALITY CHECK: TRANSFORMED  
# Checking normality of residuals in log-transformed data
# W = 0.78719, p-value < 2.2e-16

shapiro.test(resid(chl.full.B))
```
```{r chl qqnorm B}

# Transformed data: QQPlot 
qqnorm(resid(chl.full.B))

```

 
## Reduced Models  

- work bottom to top approach  
- drop the interaction  
- keep main factors  
- always keep “Treatment” and “(1|ID)” in all models  
- our goal is to find out if any of these factors improve model fit →   
- does it account for more of the variance in our response trait?  
- if it doesn’t, the chi square value will NOT be significant.  
- the factor we drop from the reduced model is the factor of focus. We want to know if adding this factor accounts for MORE variance than the treatment alone. A.K.A. we are investigating the effect of the factor we drop.  



```{r chl reduced populationxdrought}

# 1: (Population:Treatment): chisq 6.5799

chl.red.1 = lmer(Chlorophyll ~ Treatment + 
                   (1|ID) + 
                   (1|Population) + 
                   (1|Week) + 
                   (1|Population:Treatment), 
                 data = functional.phase.data, 
                 na.action = na.omit)

Anova(chl.red.1)

```

```{r chl reduced droughtxweek}

# 2: (Treatment:Time): chisq 4.1081

chl.red.2 = lmer(Chlorophyll ~ Treatment + 
                   (1|ID) + 
                   (1|Population) + 
                   (1|Week) + 
                   (1|Treatment:Week), 
                 data = functional.phase.data,  
                 na.action = na.omit)

Anova(chl.red.2)

```
- working top bottom approach, now  
- removed all the interactions and keep the main factors  
- further reduce the model by dropping a factor each model  

```{r}

# 3: no interactions: chisq 6.58

chl.red.3 = lmer(Chlorophyll ~ Treatment + 
                     (1|ID) + 
                     (1|Population) + 
                     (1|Week), 
                   data = functional.phase.data, 
                   na.action = na.omit)
Anova(chl.red.3)
```
- test the effect of week on the variance by dropping it in the reduced model and keeping population.    

```{r}

# 4: effect of week: chisq 6.0626

chl.red.4 = lmer(Chlorophyll ~ Treatment + 
                     (1|ID) + 
                     (1|Population), 
                   data = functional.phase.data, 
                   na.action = na.omit)
Anova(chl.red.4)
```
- test the effect of population on the variance by dropping it in the reduced model and keeping week.    
```{r}

# 5: effect of population: chisq 5.4025

chl.red.5 = lmer(Chlorophyll ~ Treatment + 
                         (1|ID) + 
                         (1|Week), 
                       data = functional.phase.data, 
                       na.action = na.omit)
Anova(chl.red.5)
```
## Comparison Models  
- the question: does adding the interaction/factors improve the model fit?
- comp.model.1 = anova(full.model, red.model.1) # test treatment x time interaction  
- comp.model.2 = anova(full.model, red.model.2) # test treatment x population interaction  
- comp.model.3 = anova(red.model.3, red.model.4) # test effect of week  
- comp.model.4= anova(red.model.3, red.model.5) # test effect of population    

```{r}

# 1: Test (treatment:time) interaction: chisq 2.8486

anova(chl.full.A, chl.red.1)
```

```{r}

# 2: Test (treatment:population) interaction: chisq 0

anova(chl.full.A, chl.red.2)
```

```{r}

# 3: Test the effect of week: chisq 1.226

anova(chl.red.3, chl.red.4)
```

```{r}

# 4: Test the effect of population: chisq 26.032

anova(chl.red.3, chl.red.5)
```









