---
title: '**Week 06B: Introduction to RMarkdown**'
author: "Kauanoe Greene"
date: "2024-10-08"
output: 
html_document:
toc: TRUE
toc_float: TRUE

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      fig.path = "../Output/")
```


# **Introduction**

Introduction to RMarkdown using PalmerPenguins dataset!

Datasets included:  
- *Palmer Penguins Dataset*    
- *Dr. Nyssa's Chemical Oceanography Dataset*  

<!--
Homework:
- create at least one table
- create one figure
- knit the file
- use any device you preference (pdf, html...)
- headings with clear notes
- practice bold/italics/lists
- save outputs-->


# **Packages**

```{r}

# install.packages("kableExtra")
# install.packages("ymlthis")
# install.packages("gtextras")
# install.packages("DT")
# install.packages("broom")

```
# **Libraries**

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(lubridate)
library(ggthemes)
library(ggplot2) 
library(palmerpenguins) # dataset
library(kableExtra) # kable tables
library(pixiedust)
library(DT)
library(formattable)

view(penguins)
```

<!-- check out the data -->

```{r}
glimpse(penguins)
```
# **Data**

Data section includes data upload and data analyses of penguin and chemical datasets.

#### *Palmer Penguins*

<!-- pivot_longer so we can facet wrap and plot! -->

```{r}
penguins_long <- penguins %>% 
  pivot_longer(cols = bill_length_mm:body_mass_g, 
               names_to = "Parameters",
               values_to = "Values") %>% 
  drop_na()
  
view(penguins_long)
```

#### *Chemical Data*

```{r data analyses}

# read in cond data!
# read in depth data!
# check it out :)

CondData<- read.csv(here("Week_06", "Data", "CondData.csv"))
DepthData<- read.csv(here("Week_06", "Data", "DepthData.csv")) 

glimpse(CondData) 
glimpse(DepthData) 

# formatting dataset 1: cond data
# modify cond dataset so the datetime is in character format and matches depth dataset.
# modify column name - date column to be "datetime" column
# round the datetime format to round up to nearest 10 secs.
# drop the NAs
# select which parameters you want to include in dataset.
# always check out your dataset after pipeing!

CondData_datetime<-CondData %>% 
  mutate(datetime = mdy_hm(date), 
         datetime=round_date(datetime, "10 secs")) %>% 
  drop_na() %>% 
  select(datetime, Temperature, Serial, Salinity) 

view(CondData_datetime) 


# formatting dataset 2: depth data

DepthData_datetime<-DepthData %>%
  mutate(datetime = mdy_hm(date), 
         datetime=round_date(datetime, "10 secs")) %>% 
  drop_na() %>% 
  select(datetime, AbsPressure, Depth)

view(DepthData_datetime) 

# always check out your data after pipeing!
# lubridate func task! join datasets 1 and dataset 2!
# used inner_join so there are no NAs.


datajoin<- inner_join(CondData_datetime, DepthData_datetime) 

# stats task! calculate the mean values at each minute for depth, temp, and salinity!
# time to calculate the stats
# choose which parameters you want to keep in your dataset to summarise.
# choose which how you want to group the mean by (we want it grouped by each minute).
# calculate the mean of each numeric parameter listed and group by each minute.
# list function allows us to rename the mean columns.
# reorient for facet wrap!
# name of value column
# export as CSV.


  datajoin_mean <- datajoin %>% 
    select(datetime, Depth, Temperature, Salinity) %>% 
    group_by(datetime) %>% 
    summarise(across(where(is.numeric), 
                   list(mean =mean))) %>% 
    pivot_longer(cols = Depth_mean:Salinity_mean, 
                 names_to = "Parameters", 
                 values_to = "Values") %>% 
    write_csv(here("Week_06", "Output", "Week_06_hw.summary.csv")) 
  
view(datajoin_mean) 


```

# **Plot**

Plot section includes practice penguin and chemical plots.

#### *Palmer Penguins*

```{r penguinplot, fig.align='center', fig.cap="This is a boxplot of penguin traits"}
ggplot(data = penguins_long, 
       mapping = aes(x = island, 
                     y = Values, 
                     fill = species)) + 
  geom_boxplot() + 
  facet_wrap(vars(Parameters), scales = "free") + 
  labs(title = "Species Trait Variability Across Islands", 
       subtitle = "Penguin species trait variability across Biscoe, Dream, and Torgersen Islands",
       x = "Parameters",
       y = "Values") + 
  scale_color_viridis_d() + 
  theme_classic()
```

#### *Chemical Data*

```{r chemicalplot, fig.align='center', fig.cap="This is a boxplot of chemical oceanography data"}

# plot data!
# identify the data you want to plot
# x axis variable
# y values!
# each parameter will be a different color.
# line graph!
# face wrap! plot multiple parameters along one axis.
# plot tile
# x axis title
# y axis title
# color scheme

ggplot(data = datajoin_mean, 
       mapping = aes(x = datetime, 
                     y = Values, 
                     color = Parameters)) + 
  geom_line() + 
  facet_wrap(vars(Parameters), scales = "free") + 
  labs(title = "Variability in Depth, Salinity and Temperature Over Time", 
       x = "Time", 
       y = "Mean Values") + 
  scale_color_viridis_d() 

```

# **Table**

Table section include practice tables using kabel() function with both penguin and chemical datasets.
#### *Palmer Penguins*

<!--Practice table with kable!-->
<!-- Practicing making my own kable! -->

```{r}
penguins_kbl <- penguins %>% 
  pivot_longer(cols = bill_length_mm:body_mass_g, 
               names_to = "Parameters",
               values_to = "Values") %>% 
  drop_na() %>% 
  group_by(species, Parameters) %>% 
  summarise(mean_values = mean(Values))

view(penguins_kbl)

penguins_kbl %>% 
    kbl() %>% 
    kable_classic() %>% 
    kable_styling(full_width = TRUE)
  
  

```

#### *Chemical Data*

```{r}

# using a new package (formattable) to design a table using the chemical data.

datajoin_mean %>% 
    kbl() %>% 
    kable_classic() %>% 
    kable_styling(full_width = TRUE)

```

